---
title: "Graphs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Load in the libraries and the data
```{r}
library(ggplot2)
library(knitr)
titanic <- read.csv("titanic.csv")
names(titanic) <- tolower(names(titanic))
```

### Create a new column for family size
# sibsp: The number of siblings or spouse of a person onboard
# parch: Number of Parents/Children Aboard
```{r}
titanic$family_size <- titanic$sibsp + titanic$parch + 1
```

### Q1 Bar charts

#### a. Gender of the passengers onboard
```{r}
ggplot(titanic, aes(x = sex, fill = sex)) + 
  geom_bar() + 
  labs(title = "Gender of the passengers onboard")
```

#### b. Ticket class of the passengers onboard

```{r}
library(dplyr)
titanic %>%  
  mutate(pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
  ggplot(aes(x = pclass, fill = pclass)) + 
    geom_bar() + 
    ggtitle("Ticket class of the passengers onboard") + 
    xlab("ticket class") +
    theme(legend.position="none")

```

#### c. Survival of the passengers onboard
```{r} 
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  ggplot(aes(x = survived, fill = survived)) + 
    geom_bar() + 
    ggtitle("Survival of the passengers onboard") + 
    theme(legend.position="none")
```

### Q2 Age of the passengers onboard

#### a. Histogram 
```{r}
ggplot(titanic, aes(x = age, y=..density..)) + 
  geom_histogram(fill = "steelblue", bins = 10, na.rm = TRUE) + 
  ggtitle("Distribution of age of the passengers onboard") 
```

#### b. Boxplot (group by ticket class)
```{r}
titanic %>%
  mutate(pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
ggplot(aes(x = pclass, y = age, fill = pclass)) + 
  geom_boxplot(na.rm = TRUE) + 
  ggtitle("Age of the passengers onboard") + 
  xlab("ticket class") + 
  theme(legend.position="none")
```

#### c. Boxplot (group by survival)
```{r}
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  ggplot(aes(x = survived, y = age, fill = survived)) + 
    geom_boxplot(na.rm = TRUE) + 
    ggtitle("Age of the passengers onboard") + 
    theme(legend.position="none")
```


### Q3 Ticket fare of the passengers onboard

```{r}
ggplot(titanic, aes(x = fare, y= ..density..)) + 
  geom_histogram(fill = "steelblue", bins = 30, na.rm = TRUE) + 
  ggtitle("Ticket fare of the passengers onboard") 
```

```{r}
paid_table <- table(titanic$fare != 0)
names(paid_table) <- c("not pay", "paid")
paid_table
```

### Q4 Family size per ticket class

```{r}
titanic %>%
  mutate(Pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
  ggplot(aes(x = family_size, y = ..density.., fill = pclass)) +
  geom_histogram(bins = 11) +
  facet_grid(pclass ~ ., scales = "free")  + 
  guides(fill=FALSE) +
  ggtitle("Family size per ticket class")
```

### Q5 Stacked bar charts to describe the survival rate per gender and ticket class

#### Survival rate per gender

```{r}
titanic %>%
    mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
    ggplot(aes(fill = survived, x = sex)) + 
    geom_bar(position ="stack") +
    ggtitle("Survival by gender") + 
    xlab("gender") + 
    guides(fill=guide_legend("survival"))
```

#### Survival rate per ticket class

```{r}
titanic %>%
    mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
    mutate(pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
    ggplot(aes(fill = survived, x = pclass)) + 
    geom_bar(position ="stack") +
    ggtitle("Survival by ticket class") + 
    xlab("ticket class") + 
    guides(fill=guide_legend("survival"))
```

### Q6. Violin chart describing the survival rate per age and gender 
```{r}
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  ggplot(aes(x = sex, y = age, fill = survived)) + 
    geom_violin(na.rm = TRUE)
```

### Q7. Violin chart describing the survival rate per age and ticket class
```{r}
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  mutate(pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
  ggplot(aes(x = pclass, y = age, fill = survived)) + 
    geom_violin(na.rm = TRUE) + 
    xlab("ticket class")
```
### Additional Question 1. Scatter plot of Age vs Fare, colored by Survival
```{r}
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  ggplot(aes(x = age, y = fare, color = survived)) +
    geom_point(na.rm = TRUE, alpha = 0.6) +
    ggtitle("Scatter plot of Age vs Fare, colored by Survival") +
    xlab("Age") +
    ylab("Fare") +
    guides(color = guide_legend("Survival"))
```

### Additional Question 2. Faceted bar chart of gender distribution by ticket class
```{r}
titanic %>%
  mutate(Pclass = factor(x = pclass, levels = c(1,2,3), labels = c("1st", "2nd", "3rd"))) %>%
  ggplot(aes(x = sex, fill = sex)) +
    geom_bar() +
    facet_wrap(~ pclass) +
    ggtitle("Gender distribution by ticket class") +
    xlab("Gender") +
    ylab("Count") +
    guides(fill = FALSE)
```

### Additional Question 3. Density plot of Age grouped by Survival
```{r}
titanic %>%
  mutate(survived = factor(x = survived, levels = c(0,1), labels = c("No", "Yes"))) %>%
  ggplot(aes(x = age, fill = survived)) +
    geom_density(alpha = 0.6, na.rm = TRUE) +
    ggtitle("Density plot of Age grouped by Survival") +
    xlab("Age") +
    guides(fill = guide_legend("Survival"))
```
